---
title: "Buffer Analysis"
author: "Angel Bogdanov Grigorov and Adela Sobotkova"
date: "Created 19 May 2025, updated `r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    highlight: tango
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
```

In this file we demonstrate the buffer analysis representing the straight distances from the respective cities.
Acording to Idrisi Aqranus is six days' march from Istibuni (modern-day Ihtiman), 40 miles from Farui (modern-day Stara Zagora), and four days' march from Lufisa (modern-day Lovech). It is generally accepted that 1 mile according to al-Idrisi is approximately 1.555 kilometers, and a day’s march is around 23–25 miles. Using those equations we can translate into modern distance units: 223.9 km from Istibuni, 62.2 km from Farui, and 149.2 km from Lufisa, taking the average of 23-25, which is 24.
In this case, using the buffers with the respective distances Kran should ideally be located in their intersection.

# First, loda data

```{r i}

source("LoadData.R")

```

```{r ii}

# Plot the loaded data

plot(Territory); # The territory
plot(Locations$geometry, cex=1, pch = 19, col="yellow", add=TRUE) # The locations (Lufisa, Farui, Istibuni)

```

# The Analysis

```{r iii}
# 149.2 km from Lufisa

LufisaBuff <- st_buffer(Locations[1,], dist = 149200)

# 62.2 km from Farui

FaruiBuff <- st_buffer(Locations[2,], dist = 62200)

# and 223.9 km from Istibuni

IstibuniBuff <- st_buffer(Locations[3,], dist = 223900)

```

```{r iv}
# Plot the results
options(scipen =999)
plot(Territory); # The territory
plot(Locations$geometry, cex=2, pch = 19, col=c("yellow","red", "green"), add=TRUE); # The locations (Lufisa, Farui, Istibuni)
plot(LufisaBuff, col = NA, border = "yellow", lwd = 2, add = TRUE); 
plot(FaruiBuff, col = NA, border = "red", lwd = 2, add = TRUE);
plot(IstibuniBuff, col = NA, border = "green", lwd = 2, add = TRUE)

```

# Test with 23 miles per a day

All the results seems to be out of the Kazanlak Valley. Is this a mistake of the method we use or of al-Idrisi. What if the one days march is not 24 miles as we assumed but 23. So we repeated the analysis with 23.
Because the distances from Farui are mentioned in miles, it won't be necessary to repeat this calculations. It will be done only for Istibuni and Lufisa.
In this case, the distances from 
Istibuni 6 day's march x 23 miles = 138 miles x 1,555 km per a mile should be 214,5 km
Lufisa 4 day's march x 23 miles = 92 miles x 1,555 km per a mile should be 143 km

```{r v}
# 143 km from Lufisa

LufisaBuff143km <- st_buffer(Locations[1,], dist = 143000)

# and 214.5 km from Istibuni

IstibuniBuff214.5km <- st_buffer(Locations[3,], dist = 214500)

```

```{r vi}
# Plot the results
options(scipen =999)
plot(Territory); # The territory
plot(Locations$geometry, cex=2, pch = 19, col=c("yellow","red", "green"), add=TRUE); # The locations (Lufisa, Farui, Istibuni)
plot(LufisaBuff143km, col = NA, border = "yellow", lwd = 2, add = TRUE); 
plot(FaruiBuff, col = NA, border = "red", lwd = 2, add = TRUE);
plot(IstibuniBuff214.5km, col = NA, border = "green", lwd = 2, add = TRUE)

```











