---
title: "Buffer Analysis"
author: "Angel Bogdanov Grigorov and Adela Sobotkova"
date: "Created 19 May 2025, updated `r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    highlight: tango
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
```

In this file we demonstrate the buffer analysis representing the straight distances from the respective cities.

According to al-Idrisi, Aqranus was six marches (marāḥil) from Istibuni (modern-day Ihtiman, Western Bulgaria), 40 miles (mīlān) from Farui (modern-day Stara Zagora, Central Bulgaria), and four marches (marāḥil) from Lufisa (modern-day Lovech, Northern Bulgaria)

Al-Idrisi measures distances using three main units: miles (sg. mīl, pl. mīlan or amyal, ), days of march (sg. yawm, pl. ayyām), and marches (sg. marḥalah, pl. marāḥil). Typically, he expresses short distances in miles and longer ones in days, although he is not entirely consistent. The exact value of these units is debated. It is generally accepted that one mile equals approximately 1,555 meters, a day’s march ranges between 23 and 25 miles, and a day of intensive travel can extend to 30–36 miles. The one-day trip measure varies depending on terrain and travel conditions. A marḥalah refers to the distance between two postal stations (manzil). As a term, it originates from the travels of caravans, typically covering about 7–8 hours of travel. In many translations of al-Idrisi’s work, marḥalah is rendered as “a day’s march”

In our analysis, we adopt Nedkov and others' interpretation for one mile of al-Idrisi as corresponding to approximately 1,555 meters and one marḥalah covering about 7–8 hours of travel (Недков 1960, 17). We use Murrieta-Flores’ average 4.9 km per hour for 20–60 year old men carrying nothing on his back (Murrieta-Flores 2010, 253). Applying these conversions, we derive the following modern equivalents:

6 marches from Istibuni (Ihtiman) ≈ 7 hours × 4.9 km/h ≈ 34.3 km.
6 marches × 34.3 km ≈ 205.8 km 

4 marches from Lufisa (Lovech) ≈ 7 hours × 4.9 km/h ≈ 34.3 km
4 marches × 34.3 km ≈ 137.2 km

40 miles from Farui (Stara Zagora) ≈ 40 miles × 1,555 km per mile ≈ 62.2 km

In this case, using the buffers with the respective distances Kran should ideally be located in their intersection.

# First, loda data

```{r i}

source("LoadData.R")

```

```{r ii}

# Plot the loaded data

plot(Territory); # The territory
plot(Locations$geometry, cex=1, pch = 19, col="yellow", add=TRUE) # The locations (Lufisa, Farui, Istibuni)

```

# The Analysis

```{r iii}
# 137.2 km from Lufisa

LufisaBuff <- st_buffer(Locations[1,], dist = 137200)

# 62.2 km from Farui

FaruiBuff <- st_buffer(Locations[2,], dist = 62200)

# and 205.8 km from Istibuni

IstibuniBuff <- st_buffer(Locations[3,], dist = 205800)

```



```{r iv}
# Plot the results
options(scipen =999)
plot(Territory); # The territory
plot(Locations$geometry, cex=2, pch = 19, col=c("yellow","red", "green"), add=TRUE); # The locations (Lufisa, Farui, Istibuni)
plot(LufisaBuff, col = NA, border = "yellow", lwd = 2, add = TRUE); 
plot(FaruiBuff, col = NA, border = "red", lwd = 2, add = TRUE);
plot(IstibuniBuff, col = NA, border = "green", lwd = 2, add = TRUE)

```

## Adela's test for max and min speed


```{r}
library(sf)

# Parameters (set as needed!)
arab_mile_in_feet <- 6474
feet_to_mile <- arab_mile_in_feet * 0.3
mile_to_km <- 1.555
max_miles_per_day <- 15  # 15 miles/per day == 23 km/day
min_miles_per_day <- 10

# Max distance (same as your original)
distL_max  <- 4 * max_miles_per_day * mile_to_km
distF_max  <- 40 * mile_to_km
distI_max  <- 6 * max_miles_per_day * mile_to_km

# Min distance (lower daily walking range)
distL_min  <- 4 * min_miles_per_day * mile_to_km
distF_min  <- 40 * (min_miles_per_day / max_miles_per_day) * mile_to_km  # proportional reduction
distI_min  <- 6 * min_miles_per_day * mile_to_km

# Create buffers in meters (multiply km by 1000)
# Lufisa
LufisaBuff_max <- st_buffer(Locations[1,], dist = distL_max * 1000)
LufisaBuff_min <- st_buffer(Locations[1,], dist = distL_min * 1000)
LufisaDonut <- st_difference(LufisaBuff_max, LufisaBuff_min)

# Farui
FaruiBuff_max <- st_buffer(Locations[2,], dist = distF_max * 1000)
FaruiBuff_min <- st_buffer(Locations[2,], dist = distF_min * 1000)
FaruiDonut <- st_difference(FaruiBuff_max, FaruiBuff_min)

# Istibuni
IstibuniBuff_max <- st_buffer(Locations[3,], dist = distI_max * 1000)
IstibuniBuff_min <- st_buffer(Locations[3,], dist = distI_min * 1000)
IstibuniDonut <- st_difference(IstibuniBuff_max, IstibuniBuff_min)

```
```{r}
# Combine all donut zones
Donuts <- rbind(
  LufisaDonut %>% mutate(Source = "Lufisa", color = "yellow"),
  FaruiDonut %>% mutate(Source = "Farui", color = "red"),
  IstibuniDonut %>% mutate(Source = "Istibuni", color = "green")
)
```



```{r}

# Plot raster background (elevation > 0)
plot(Territory, col = terrain.colors(25), main = "Donut Buffers by Site: Uncertainty in Daily March Distance (20-25)")

# Add donuts
plot(Donuts["geometry"], border = Donuts$color, col = adjustcolor(Donuts$color, alpha.f = 0.4), add = TRUE, lwd = 2)

# Add locations
plot(Locations$geometry, pch = 21, bg = "black", col = c("yellow", "red", "green"), cex = 1.5, add = TRUE)

```

```{r}
# Create interactive map
mapview(Donuts, zcol = "color", layer.name = "Walking Range Donuts", alpha.regions = 0.4, lwd = 2) +
  mapview(Locations, col.regions = c("yellow", "red", "green"), layer.name = "Locations", cex = 3)
```

# References

Murrieta-Flores, Patricia. 2010. ‘Travelling in a Prehistoric Landscape: Exploring the Influences That Shaped Human Movement’. In Making History Interactive. Computer Applications and Quantitative Methods in Archaeology (CAA), edited by Bernard Fisher, Jane Crawford, and David Koller, 249–67. BAR International. Oxford: Archaeopress. https://doi.org/10.15496/PUBLIKATION-3314.

Недков, Борис. 1960. ‘Въведение’. In България и съседните ѝ земи през XII в. според Идриси, edited by Борис Недков, 9–29. София: Наука и изкуство.

