---
title: "Walking Isochrones Analysis"
author: "Angel Bogdanov Grigorov and Adela Sobotkova"
date: "Created 10 Oct. 2024, updated `r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    highlight: tango
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
```

In these file we demonstrate the Walking Isochrones analysis. The text of al-Idrisi states that Aqranus is six days' march from Istibuni (modern-day Ihtiman), 40 miles from Farui (modern-day Stara Zagora), and four days' march from Lufisa (modern-day Lovech). One mile according to al-Idrisi is approximately 1.555 kilometers, and a day’s march is around 23–25 miles. Using those equations we can translate into modern distance units: 223.9 km from Istibuni, 62.2 km from Farui, and 149.2 km from Lufisa, taking the average of 23-25, which is 24. It is generally accepted it takes circa 12 minutes to walk 1 km or 1 hour to walk 5 km. We transformed the distances into walking hours:
6 days' march from Istibuni ≈ 223.9 km / 5 km per h ≈ 44.7 walking h
40 miles from Farui ≈ 62.2 km / 5 km per h ≈ 12.4 walking h
4 days' march from Lufisa ≈ 149.2 km / 5 km per h ≈ 29.8 walking h

# First, load data and library

```{r i}

source("LoadData.R")
library(movecost)
library(raster)
```

# Aggregate the Territory

```{r ii}

# If the file is too heavy it may be necessary to aggregate the DEM you are using:

# TerAgg2 <- raster::aggregate(Territory, fact = 2, fun = mean, na.rm = T) # Aggregate double the size

# TerAgg3 <- raster::aggregate(Territory, fact = 3, fun = mean, na.rm = T) # Aggregate triple the size

# ...

TerAgg10 <- aggregate(Territory, fact = 10, fun = mean, na.rm = T) # Aggregate the Territory 10 times

```

# Plot the data

```{r iii}

# Let's visualise

plot(TerAgg10); plot(Locations$geometry, cex=1, pch = 19, col=c("yellow","red", "green"), add=TRUE)

```

# The Analysis

```{r iv}

# Because we run the function in minutes, the calculations are as follows:
# If Lufisa is 149.2 from Aqranus it is  149.2 km / 5 km per h = 29.8 hours, which represented as minutes should be 1790.4 min (149.8 km х 12 min = 1790.4).
# Respectively for the other cities down

LufisaIso <- movebound(dtm=raster(Territory),  # use original raster as it's already lo-res at 30s
                       origin=as(Locations[1, ], "Spatial"), 
                       cont.value = 1790, time = "m", move=8, # We round 1790.4 to 1790
                       funct = "t", cogn.slp = TRUE, export = FALSE) # export fails so do it manually 


# Visual check and export isoline vector
LufisaIso$isolines %>% 
  st_as_sf() %>% 
  st_set_crs(32635) %>% 
  mapview() # view and save as shp if you need to
```

```{r v}

# Farui

FaruiIso <- movebound(dtm=raster(Territory),  # use original raster as it's already lo-res at 30s
                      origin=as(Locations[2, ], "Spatial"), 
                      cont.value = 746, time = "m", move=8, funct = "t", # We round 746.4 to 746
                      cogn.slp = TRUE, export = FALSE) 

# Visual check and export isoline vector
FaruiIso$isolines %>% 
  st_as_sf() %>% 
  st_set_crs(32635) %>% 
  mapview() # view and save as shp if you need to
```

```{r vi}

# Istibuni

IstibuniIso <- movebound(dtm=raster(Territory),  # use original raster as it's already lo-res at 30s
                         origin=as(Locations[3, ], "Spatial"), cont.value = 2686, # We round 2686.8 to 2686
                         time = "m", move=8, funct = "t", 
                         cogn.slp = TRUE, export = FALSE) 

# Visual check and export isoline vector
IstibuniIso$isolines %>% 
  st_as_sf() %>% 
  st_set_crs(32635) %>% 
  mapview() # view and save as shp if you need to

```

```{r vii}
# Plot the results

plot(Territory); # The territory
plot(Locations$geometry, cex=2, pch = 19, col=c("yellow","red", "green"), add=TRUE); # The locations (Lufisa, Farui, Istibuni)
plot(LufisaIso$isolines, col="yellow", lwd=2,add = TRUE); 
plot(FaruiIso$isolines, col="red", lwd =2,add = TRUE);
plot(IstibuniIso$isolines, col="green", lwd=2, add = TRUE)

```

# Test with 23 miles per a day

All the results seems to be out of the Kazanlak Valley. Is this a mistake of the method we use or of al-Idrisi. What if the one days march is not 24 miles as we assumed but 23. So we repeated the analysis with 23.
Because the distances from Farui are mentioned in miles, it won't be necessary to repeat this calculations. It will be done only for Istibuni and Lufisa.
In this case, the distances from 
Istibuni should be 214,5 km / 5 km = 42,9 walking h (or 2,574 mins)
Lufisa should be 143 km  / 5 km = 28,6 walking h (1,716 mins)

```{r viii}

# Lufisa

LufisaIso1716 <- movebound(dtm=raster(Territory),  # use original raster as it's already lo-res at 30s
                       origin=as(Locations[1, ], "Spatial"), 
                       cont.value = 1716, time = "m", move=8,
                       funct = "t", cogn.slp = TRUE, export = FALSE) # export fails so do it manually 


# Visual check and export isoline vector
LufisaIso1716$isolines %>% 
  st_as_sf() %>% 
  st_set_crs(32635) %>% 
  mapview() # view and save as shp if you need to
```


```{r ix}

# Istibuni

IstibuniIso2574 <- movebound(dtm=raster(Territory),  # use original raster as it's already lo-res at 30s
                         origin=as(Locations[3, ], "Spatial"), cont.value = 2574,
                         time = "m", move=8, funct = "t", 
                         cogn.slp = TRUE, export = FALSE) 

# Visual check and export isoline vector
IstibuniIso$isolines %>% 
  st_as_sf() %>% 
  st_set_crs(32635) %>% 
  mapview() # view and save as shp if you need to

```


```{r x}
# Plot the results

plot(Territory); # The territory
plot(Locations$geometry, cex=2, pch = 19, col=c("yellow","red", "green"), add=TRUE); # The locations (Lufisa, Farui, Istibuni)
plot(LufisaIso1716$isolines, col="yellow", lwd=2,add = TRUE); 
plot(FaruiIso$isolines, col="red", lwd =2,add = TRUE);
plot(IstibuniIso2574$isolines, col="green", lwd=2, add = TRUE)

```







