---
title: "Walking Isochrones Analysis"
author: "Angel Bogdanov Grigorov and Adela Sobotkova"
date: "Created 10 Oct. 2024, updated `r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    highlight: tango
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
```

In these file we demonstrate the Walking Isochrones analysis. 

According to al-Idrisi, Aqranus was six marāḥil from Istibuni (modern-day Ihtiman, Western Bulgaria), 40 mīlān from Farui (modern-day Stara Zagora, Central Bulgaria), and four marāḥil from Lufisa (modern-day Lovech, Northern Bulgaria)

Typically, he expresses short distances in miles and longer ones in days, although he is not entirely consistent. The exact value of these units is debated. It is generally accepted that one mile equals approximately 1,555 meters. A day of walk is usually translated as a distance, stretching from a moderate distance of 23 - 25 miles to intensive travel of 30–36 miles. A marching day originates from the travel of caravans. It is counted not in miles but rather as time, specifically 7–8 hours of travel.  Sometimes translators of al-Idrisi’s work do not differentiate between marḥalah and yawm, rendering both as distances equivalent to the “day of walk” (Недков 1960, 17; Maqbul 1992, 160).

In our analysis, we adopt Nedkov and others' interpretation for one mile of al-Idrisi as corresponding to approximately 1,555 meters and one marḥalah covering about 7–8 hours of travel (Недков 1960, 17). We use Murrieta-Flores’ average 4.9 km per hour for 20–60 year old men carrying nothing on his back (Murrieta-Flores 2010, 253). Applying these conversions, we derive the following modern equivalents:

6 marāḥil from Istibuni (Ihtiman) ≈ 6  ×  7 hours × 4.9 km/h ≈ 205.8 km 

4 marāḥil from Lufisa (Lovech) ≈ 4 × 7 hours × 4.9 km/h ≈ 137.2 km

40 mīlan from Farui (Stara Zagora) ≈ 40 miles × 1,555 km per mile ≈ 62.2 km

In order to account for the friction of traversing the mountainous terrain in central Bulgaria, we calculate the cost of travel between the neighboring towns next. We have to convert al-Idrisi’s reported distances into estimated walking times. Al-Idrisi refers to ‘march’, which we translate into the walking speed of 4.9 km per hour, which is the average for a male pedestrian between 20 and 60 years of age (Murrieta-Flores 2010, 253). The following inputs represent the maximum time spent in marching to Aqranus from each of the three cities:

6 marāḥil from Istibuni ≈ 6 x 7 h ≈ 42 walking hours

4 marāḥil from Lufisa ≈ 4 x 7 h ≈ 28 walking hours

40 mīlan from Farui ≈ 62.2 km ÷ 4.9 km/h ≈ 12.6 walking hours

These values are used to generate walking-time isochrones around each reference city, with the boundary representing maximum reachable distance. 


# First, load data and library

```{r i}

source("LoadData.R")
library(movecost)
library(raster)
```

# Aggregate the Territory

```{r ii}

# If the file is too heavy it may be necessary to aggregate the DEM you are using:

# TerAgg2 <- raster::aggregate(Territory, fact = 2, fun = mean, na.rm = T) # Aggregate double the size

# TerAgg3 <- raster::aggregate(Territory, fact = 3, fun = mean, na.rm = T) # Aggregate triple the size

# ...

TerAgg10 <- aggregate(Territory, fact = 10, fun = mean, na.rm = T) # Aggregate the Territory 10 times

```

# Plot the data

```{r iii}

# Let's visualise

plot(Territory); plot(Locations$geometry, cex=1, pch = 19, col=c("yellow","red", "green"), add=TRUE)
plot(TerAgg10); plot(Locations$geometry, cex=1, pch = 19, col=c("yellow","red", "green"), add=TRUE)

```

# The Analysis

```{r iv}

# Because we run the function in minutes, the calculations are as follows:
# If Aqranus is 4 marahil from Lufisa (Lovech) ≈ 4 marahil x 7 h ≈ 28 h x 60 min = 1680 min.

LufisaIso <- movebound(dtm=raster(Territory),  # use original raster as it's already lo-res at 30s
                       origin=as(Locations[1, ], "Spatial"),
                         cont.value = 1680, time = "m", move=8, # using 4 marches * 7 hrs *60 mins
                       funct = "t", cogn.slp = TRUE, N = 1.02, # Set N = 1.02 adjust speed to ~4.9 km/h
                       export = FALSE) # Export fails so do it manually

# Visual check and export isoline vector
LufisaIso$isolines %>% 
  st_as_sf() %>% 
  st_set_crs(32635) %>% 
  mapview() # view and save as shp if you need to
```

```{r v}

# If Aqranus is 40 miles from Farui (Stara Zagora) ≈ 62.2 km ÷ 4.9 km/h ≈ 12,6 h x 60 min ≈ 756 min

FaruiIso <- movebound(dtm=raster(Territory),  # use original raster as it's already lo-res at 30s
                      origin=as(Locations[2, ], "Spatial"), 
                      cont.value = 756, time = "m", move=8, funct = "t",
                      cogn.slp = TRUE, N = 1.02, # Set N = 1.02 adjust speed to ~4.9 km/h
                      export = FALSE) # # Export fails so do it manually

# Visual check and export isoline vector
FaruiIso$isolines %>% 
  st_as_sf() %>% 
  st_set_crs(32635) %>% 
  mapview() # view and save as shp if you need to
```

```{r vi}

# If Aqranus is 6 marahil from Istibuni (Ihtiman) ≈ 6 marahil x 7 h ≈ 42 h x 60 min ≈ 2520 min

IstibuniIso <- movebound(dtm=raster(Territory),  # use original raster as it's already lo-res at 30s
                         origin=as(Locations[3, ], "Spatial"),
                         cont.value = 2520, #  using 7 days * 6 hrs *60 mins
                         time = "m", move=8, funct = "t", 
                         cogn.slp = TRUE, N = 1.02, # Set N = 1.02 adjust speed to ~4.9 km/h
                         export = FALSE) # # Export fails so do it manually 

# Visual check and export isoline vector
IstibuniIso$isolines %>% 
  st_as_sf() %>% 
  st_set_crs(32635) %>% 
  mapview() # view and save as shp if you need to

```

```{r vii}
# Plot the results

plot(Territory); # The territory
plot(Locations$geometry, cex=2, pch = 19, col=c("yellow","red", "green"), add=TRUE); # The locations (Lufisa, Farui, Istibuni)
plot(LufisaIso$isolines, col="yellow", lwd=2,add = TRUE); 
plot(FaruiIso$isolines, col="red", lwd =2,add = TRUE);
plot(IstibuniIso$isolines, col="green", lwd=2, add = TRUE)

```

# References

Maqbul, Ahmad. 1992. ‘Cartography of Al-Sharīf al-Idrīsī’. In Cartography in the Traditional Islamic and South Asian Societies, edited by Brian Harley and David Woodward, 2:156–74. The History of Cartography 1. Chicago: University of Chicago Press. https://press.uchicago.edu/books/hoc/HOC_V2_B1/Volume2_Book1.html.

Murrieta-Flores, Patricia. 2010. ‘Travelling in a Prehistoric Landscape: Exploring the Influences That Shaped Human Movement’. In Making History Interactive. Computer Applications and Quantitative Methods in Archaeology (CAA), edited by Bernard Fisher, Jane Crawford, and David Koller, 249–67. BAR International. Oxford: Archaeopress. https://doi.org/10.15496/PUBLIKATION-3314.

Недков, Борис. 1960. ‘Въведение’. In България и съседните ѝ земи през XII в. според Идриси, edited by Борис Недков, 9–29. София: Наука и изкуство.
