---
title: "Walking Isochrones Analysis"
author: "Angel Bogdanov Grigorov and Adela Sobotkova"
date: "Created 10 Oct. 2024, updated `r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    highlight: tango
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
```

In these file we demonstrate the Walking Isochrones analysis. 

Al-Idrisi measures distances using three main units: miles (sg. mīl, pl. mīlan or amyal, ), days of march (sg. yawm, pl. ayyām), and marches (sg. marḥalah, pl. marāḥil). Typically, he expresses short distances in miles and longer ones in days, although he is not entirely consistent. The exact value of these units is debated. It is generally accepted that one mile equals approximately 1,555 meters, a day’s march ranges between 23 and 25 miles, and a day of intensive travel can extend to 30–36 miles. The one-day trip measure varies depending on terrain and travel conditions. A marḥalah refers to the distance between two postal stations (manzil). As a term, it originates from the travels of caravans, typically covering about 7–8 hours of travel. In many translations of al-Idrisi’s work, marḥalah is rendered as “a day’s march”

In our analysis, we adopt Nedkov and others' interpretation for one mile of al-Idrisi as corresponding to approximately 1,555 meters and one marḥalah covering about 7–8 hours of travel (Недков 1960, 17). We use Murrieta-Flores’ average 4.9 km per hour for 20–60 year old men carrying nothing on his back (Murrieta-Flores 2010, 253). 

The following inputs represent the maximum time spent in marching to Aqranus from each of the three cities:

6 marches from Istibuni = 6 x 7 h ≈ 42 walking hours
40 miles from Farui = 62.2 km ÷ 4.9 km/h = 12.6 walking hours
4 marches from Lufisa = 4 x 7 h ≈ 28 walking hours

These values are used to generate walking-time isochrones around each reference city, with the boundary representing maximum reachable distance

# First, load data and library

```{r i}

source("LoadData.R")
library(movecost)
library(raster)
```

# Aggregate the Territory

```{r ii}

# If the file is too heavy it may be necessary to aggregate the DEM you are using:

# TerAgg2 <- raster::aggregate(Territory, fact = 2, fun = mean, na.rm = T) # Aggregate double the size

# TerAgg3 <- raster::aggregate(Territory, fact = 3, fun = mean, na.rm = T) # Aggregate triple the size

# ...

TerAgg10 <- aggregate(Territory, fact = 10, fun = mean, na.rm = T) # Aggregate the Territory 10 times

```

# Plot the data

```{r iii}

# Let's visualise

plot(Territory); plot(Locations$geometry, cex=1, pch = 19, col=c("yellow","red", "green"), add=TRUE)
plot(TerAgg10); plot(Locations$geometry, cex=1, pch = 19, col=c("yellow","red", "green"), add=TRUE)

```

# The Analysis

```{r iv}

# Because we run the function in minutes, the calculations are as follows:
# If Aqranus is 4 marches (marahil) from Lufisa (Lovech) ≈ 4 marahil x 7 h ≈ 28 h x 60 min = 1680 min.

LufisaIso <- movebound(dtm=raster(Territory),  # use original raster as it's already lo-res at 30s
                       origin=as(Locations[1, ], "Spatial"),
                         cont.value = 1680, time = "m", move=8, # using 4 marches * 7 hrs *60 mins
                       funct = "t", cogn.slp = TRUE, N = 1.02, # Set N = 1.02 adjust speed to ~4.9 km/h
                       export = FALSE) # Export fails so do it manually

# Visual check and export isoline vector
LufisaIso$isolines %>% 
  st_as_sf() %>% 
  st_set_crs(32635) %>% 
  mapview() # view and save as shp if you need to
```

```{r v}

# If Aqranus is 40 miles from Farui (Stara Zagora) ≈ 62.2 km ÷ 4.9 km/h ≈ 12,6 h x 60 min ≈ 756 min

FaruiIso <- movebound(dtm=raster(Territory),  # use original raster as it's already lo-res at 30s
                      origin=as(Locations[2, ], "Spatial"), 
                      cont.value = 756, time = "m", move=8, funct = "t",
                      cogn.slp = TRUE, N = 1.02, # Set N = 1.02 adjust speed to ~4.9 km/h
                      export = FALSE) # # Export fails so do it manually

# Visual check and export isoline vector
FaruiIso$isolines %>% 
  st_as_sf() %>% 
  st_set_crs(32635) %>% 
  mapview() # view and save as shp if you need to
```

```{r vi}

# If Aqranus is 6 marahil (marches) from Istibuni (Ihtiman) ≈ 6 marahil x 7 h ≈ 42 h x 60 min ≈ 2520 min

IstibuniIso <- movebound(dtm=raster(Territory),  # use original raster as it's already lo-res at 30s
                         origin=as(Locations[3, ], "Spatial"),
                         cont.value = 2520, #  using 7 days * 6 hrs *60 mins
                         time = "m", move=8, funct = "t", 
                         cogn.slp = TRUE, N = 1.02, # Set N = 1.02 adjust speed to ~4.9 km/h
                         export = FALSE) # # Export fails so do it manually 

# Visual check and export isoline vector
IstibuniIso$isolines %>% 
  st_as_sf() %>% 
  st_set_crs(32635) %>% 
  mapview() # view and save as shp if you need to

```

```{r vii}
# Plot the results

plot(Territory); # The territory
plot(Locations$geometry, cex=2, pch = 19, col=c("yellow","red", "green"), add=TRUE); # The locations (Lufisa, Farui, Istibuni)
plot(LufisaIso$isolines, col="yellow", lwd=2,add = TRUE); 
plot(FaruiIso$isolines, col="red", lwd =2,add = TRUE);
plot(IstibuniIso$isolines, col="green", lwd=2, add = TRUE)

```

# References

Murrieta-Flores, Patricia. 2010. ‘Travelling in a Prehistoric Landscape: Exploring the Influences That Shaped Human Movement’. In Making History Interactive. Computer Applications and Quantitative Methods in Archaeology (CAA), edited by Bernard Fisher, Jane Crawford, and David Koller, 249–67. BAR International. Oxford: Archaeopress. https://doi.org/10.15496/PUBLIKATION-3314.

Недков, Борис. 1960. ‘Въведение’. In България и съседните ѝ земи през XII в. според Идриси, edited by Борис Недков, 9–29. София: Наука и изкуство.
