---
title: "Walking Isochrones Analysis"
author: "Angel Bogdanov Grigorov"
date: "Created 10 Oct. 2024, updated `r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    highlight: tango
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
```

In these file I demonstrate the Walking Isochrones analysis. The text of al-Idrisi states that Aqranus is six days' march from Istibuni (modern-day Ihtiman), 40 miles from Farui (modern-day Stara Zagora), and four days' march from Lufisa (modern-day Lovech). One mile according to al-Idrisi is approximately one and a half kilometers, and a day’s march is around 23–25 miles. Using those equations we can translate into modern distance units: 217.7 km from Istibuni, 62.2 km from Farui, and 144.9 km from Lufisa. It is generally accepted it takes circa 12 minutes to walk 1 km or 1 hour to walk 5 km. I transformed the distances into walking hours:
6 days' march from Istibuni ≈ 217 km / 5 km per h ≈ 43 walking h
40 miles from Farui ≈ 62 km / 5 km per h ≈ 12 walking h
4 days' march from Lufisa ≈ 144 km / 5 km per h ≈ 28 walking h

# First, loda data

```{r i}

source("LoadData.R")

```

# Aggregate the Territory

```{r ii}

# If the file is too heavy it may be necesary to aggregate the DEM you are using:

# TerAgg2 <- raster::aggregate(Territory, fact = 2, fun = mean, na.rm = T) # Aggregate double the size

# TerAgg3 <- raster::aggregate(Territory, fact = 3, fun = mean, na.rm = T) # Aggregate triple the size

# ...

TerAgg10 <- raster::aggregate(Territory, fact = 10, fun = mean, na.rm = T) # Aggregate the Territory 10 times

```

# Plot the data

```{r iii}

# Let's visualise

plot(TerAgg10); plot(Locations$geometry, cex=1, pch = 19, col="red", add=TRUE)

```

# The Analysis

```{r iv}

# Because I run the function in minutes, the calculations are as follows:
# If Lufisa is 144.9 from Aqranus it is  144.9 km / 5 km per h = 28.89 hours, which represented as minutes should be 1738.8 min (144.9 km х 12 min = 1738.8).
# Respectively for the other cities down

LufisaIso <- movebound(dtm=TerAgg10, origin=as(Locations[1, ], "Spatial"), cont.value = 1739, time = "m", move=8, funct = "t", cogn.slp = TRUE, export = TRUE) 

```

```{r v}

FaruiIso <- movebound(dtm=TerAgg10, origin=as(Locations[2, ], "Spatial"), cont.value = 746, time = "m", move=8, funct = "t", cogn.slp = TRUE, export = TRUE) 

```

```{r vi}

IstibuniIso <- movebound(dtm=TerAgg10, origin=as(Locations[3, ], "Spatial"), cont.value = 2612, time = "m", move=8, funct = "t", cogn.slp = TRUE, export = TRUE) 

```

```{r vii}

# Plot the results

plot(Territory); # The territory
plot(Locations$geometry, cex=1, pch = 19, col="blue", add=TRUE); # The locations (Lufisa, Farui, Istibuni)
plot(LufisaIso$isolines, col="yellow", add = TRUE); 
plot(FaruiIso$isolines, col="red", add = TRUE);
plot(IstibuniIso$isolines, col="green", add = TRUE)

```


