---
title: "Buffer Analysis"
author: "Angel Bogdanov Grigorov"
date: "Created 19 May 2025, updated `r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    highlight: tango
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
```

In this file I demonstrate the buffer analysis representing the straight distances from the respective cities.
Acording to Idrisi Aqranus is six days' march from Istibuni (modern-day Ihtiman), 40 miles from Farui (modern-day Stara Zagora), and four days' march from Lufisa (modern-day Lovech). It is generally accepted that 1 mile according to al-Idrisi is approximately one and a half kilometers, and a day’s march is around 23–25 miles. Using those equations we can translate into modern distance units: 217.7 km from Istibuni, 62.2 km from Farui, and 144.9 km from Lufisa.
In this case, using the buffers with the respective distances Kran should ideally be located in their intersection.

# First, loda data

```{r i}

source("LoadData.R")

```

```{r ii}

# Plot the loaded data

plot(Territory); # The territory
plot(Locations$geometry, cex=1, pch = 19, col="blue", add=TRUE) # The locations (Lufisa, Farui, Istibuni)

```

# The Analysis

```{r iii}
# 144.9 km from Lufisa

LufisaBuff <- st_buffer(Locations[1,], dist = 144900)

# 62.2 km from Farui

FaruiBuff <- st_buffer(Locations[2,], dist = 62200)

# and 217.7 km from Istibuni

IstibuniBuff <- st_buffer(Locations[3,], dist = 217700)

```

```{r iV}

# Plot the results

plot(Territory); # The territory
plot(Locations$geometry, cex=1, pch = 19, col="blue", add=TRUE); # The locations (Lufisa, Farui, Istibuni)
plot(LufisaBuff, col = NA, border = "yellow", lwd = 2, add = TRUE); 
plot(FaruiBuff, col = NA, border = "red", lwd = 2, add = TRUE);
plot(IstibuniBuff, col = NA, border = "green", lwd = 2, add = TRUE)

```




















